---
title: "Experiment 3"
output: html_document
date: "2024-12-21"
---

## Children leverage the Successor Representation for reward-guided choice

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(readr)
library(ggh4x)
library(afex)
library(Rmisc)
library(dplyr)
library(kableExtra)
library(tidyr)
library(ggplot2)
library(reshape2)
library(sjPlot)
```

```{r define model functions, include=FALSE}
my_tab_model <- function(model){
  tab_model(model, 
          df.method = "wald",
          show.stat=TRUE,
          CSS = list(css.tdata = 'padding:0.1cm;padding-left:0.2cm;padding-right:0.2cm;'),
          transform = NULL,
          string.stat = "z",
          string.p = "p > |z|",
          show.re.var = FALSE,
          show.icc = FALSE,
          show.r2 = FALSE,
          wrap.labels = 100,
          show.ngroups = FALSE,
          show.obs = FALSE)
}

my_plot_model <- function(model){
  plot_model(model, 
          show.intercept = T,
          value.offset = 0.35,
          transform = NULL,
          show.values = T,) + 
  geom_hline(yintercept = 0, linetype = "dashed", alpha = .5)
}
```

### Exclusion Criteria
Participants were excluded for >20 browser interactions, >3 comprehension errors, >10 timeouts, or mean RT >5000ms 
```{r exclusion criteria}
exclusions_data <- read_csv("../data/preprocessed/excluded.csv")

criteria <- list(
  broswer_interact = function(df) df$browser_interactions > 20,
  timeouts = function(df) df$missed_trials > 10,
  comprehension_error = function(df) df$comp_retries > 3,
  mean_rt = function(df) df$avg_rt > 5000
)

criteria_counts <- data.frame(
  Criteria = character(length(criteria)),
  Count = integer(length(criteria)))

for (i in seq_along(criteria)) {
  # Get the criterion function
  criterion_name <- names(criteria)[i]
  criterion <- criteria[[i]]
  # Apply the criterion to count the number of rows
  count <- sum(criterion(exclusions_data))
  criteria_counts[i, "Criteria"] <- criterion_name
  criteria_counts[i, "Count"] <- count
  exclusions_data <- exclusions_data[!criterion(exclusions_data), ]
}

criteria_counts 
```

### Loading Full Dataset 

Reading in preprocessed dataset. Each row represents a traversal trial and its preceeding shop-only trial. All continuous variables are z-scored \

* **action1TowardsPrevEnd** (1 or 0; 1 corresponds to taking action1 towards island containing prior shop; 0 is other island).   \
* **prior_reward** (1 or 0; 1 corresponds to reward on prior shop trial). \
* **shop_neighbor_prior_reward** (1 or 0, corresponds to reward when the neighbor to the prior shop was last seen in traversal or shop-only).  \
* **prior_shop_choice** (1 or 0, 1 if prior shop was chosen on last visit to its island).   \

```{r read main dataset}
data <- read_csv("../data/preprocessed/data.csv")
scale_this <- function(x){
  (x - mean(x, na.rm=TRUE)) / sd(x, na.rm=TRUE)
}
data$shop_neighbor_prior_reward <- scale_this(data$shop_neighbor_prior_reward)
data$prior_shop_choice <- scale_this(data$prior_shop_choice)
data$prior_reward <- scale_this(data$prior_reward)
data$prior_island_reward <- scale_this(data$prior_island_reward)
data$sub <- factor(data$sub)
data$age <- scale_this(data$age)
data$categorical_age <- factor(data$categorical_age, levels = c('Children', 'Adolescents', 'Adults'))

# Filter data for analyses on children only
child_only_data <- data[data$categorical_age == 'Children',]
```

### Participants leveraged mental models to guide choice
A significant effect of **prior_reward** on island choice would suggest influence of prior shop trial on traversal trial island choice in a model-based or SR-dependent manner

```{r reward_only model}
reward_only <- glmer(action1TowardsPrevEnd ~ prior_reward * age + (prior_reward | sub), 
                     data = data, 
                     family = binomial)
my_tab_model(reward_only)
```

### Unique signatures of MB and SR-based decision strategies
* Significant effect of **prior_reward** would suggest influence of prior shop trial on traversal trial island choice in a model-based or SR-dependent manner. \
* Significant **prior_shop_choice * prior_reward** interaction indicates SR-dependent behavior, while **prior_reward * shop_neighbor_prior_reward** interaction indicates MB-behavior. \

```{r model-based/sr model, cache=TRUE}
mb_sr <- glmer(action1TowardsPrevEnd ~ prior_reward * shop_neighbor_prior_reward * prior_shop_choice * age + (prior_reward * shop_neighbor_prior_reward * prior_shop_choice | sub),
                     data = data, 
                     family = binomial)
my_tab_model(mb_sr)
```

### Refitting with child data only to see if there is still evidence for use of MB and SR strategies
```{r child only}
child_only_mb_sr <- glmer(action1TowardsPrevEnd ~ prior_reward * shop_neighbor_prior_reward * prior_shop_choice + (prior_reward * shop_neighbor_prior_reward * prior_shop_choice | sub), 
                     data = child_only_data, 
                     family = binomial) 
my_tab_model(child_only_mb_sr)
```

### Plots for SR and MB interactions

``` {r recoding variables}
# recoding variables as factors for plotting
data$prior_reward <- factor(data$prior_reward, labels=c("No Reward","Reward"))
data$shop_neighbor_prior_reward <- factor(data$shop_neighbor_prior_reward, labels=c("No Reward","Reward"))
data$prior_shop_choice <- factor(data$prior_shop_choice, labels=c("Neighboring","Observed"))
data$prior_island_reward <- factor(data$prior_island_reward, labels=c("No Reward","Reward"))
data$block_change_congruent <- factor(data$block_change_congruent, labels=c("Incongruent","Congruent"))
```

``` {r plot mb by age}
summary_data <- summarySE(data, measurevar="action1TowardsPrevEnd",
                        groupvars=c("shop_neighbor_prior_reward","prior_reward","sub","categorical_age"))
summary_data <- summarySE(data, measurevar="action1TowardsPrevEnd",
                        groupvars=c("shop_neighbor_prior_reward","prior_reward","categorical_age"))
strip <- strip_themed(background_x = elem_list_rect(fill = c("#B27EE0","#7c44ad", "#501b80")),text_x=list(element_text(colour = "white")))

p1 <- ggplot(summary_data, aes(x = shop_neighbor_prior_reward, y = action1TowardsPrevEnd, fill = prior_reward)) +
  facet_wrap2(~ categorical_age, strip=strip) +
  geom_bar(stat = "identity", position = "dodge", colour="black") +
  geom_errorbar(aes(ymin = action1TowardsPrevEnd - se, ymax = action1TowardsPrevEnd + se), 
                position = position_dodge(width = 0.9), width = 0.25) +
  labs(title = "Empirical data",
       x = "Previous reward from neighboring shop",
       y = "Prob. of choosing associated island") +
  scale_fill_manual(name = "Non-traversal reward", values=c("#7ECAC9","#325379")) +
  scale_x_discrete(guide = guide_axis(n.dodge = 2)) +
  ylim(0,.7) +
  theme_classic(base_size = 12) + 
  theme(axis.text.x=element_text(size=11)) 
ggsave("figures/experiment3_mb.png")
```

```{r test}
summary_data <- data.frame(
  shop_neighbor_prior_reward = c("No Reward", "No Reward", "Reward", "Reward"),
  action1TowardsPrevEnd = c(0.25, 0.666, 0.666, 0.666),
  prior_reward = c("No Reward", "Reward", "No Reward", "Reward")
)
p2 <- ggplot(summary_data, aes(x = shop_neighbor_prior_reward, y = action1TowardsPrevEnd, fill = prior_reward)) +
  geom_bar(stat = "identity", position = "dodge", colour="black") +
  labs(title = "Model-based prediction",
    x = "Previous reward from neighboring shop",
       y = "Prob. of choosing associated island") +
  ylim(0,.7) +
  scale_fill_manual(name = "Non-traversal reward", values=c("#7ECAC9","#325379")) +
  theme_classic(base_size = 12) + 
  theme(axis.text.x=element_text(size=11)) + 
  theme(legend.position = "none")

ggsave("figures/experiment3_mb_pred.png")

combined_plot <- p2 + plot_spacer() + p1 + plot_layout(ncol = 3, width=c(1,.2,3))
ggsave("figures/combined_plot.png", combined_plot, width = 11, height = 11/3)


summary_data <- summarySE(data, measurevar="action1TowardsPrevEnd",
                        groupvars=c("prior_reward","sub","categorical_age"))
summary_data <- summarySE(data, measurevar="action1TowardsPrevEnd",
                        groupvars=c("prior_reward","categorical_age"))
strip <- strip_themed(background_x = elem_list_rect(fill = c("#B27EE0","#7c44ad", "#501b80")),text_x=list(element_text(colour = "white")))

p1 <- ggplot(summary_data, aes(x = prior_reward, y = action1TowardsPrevEnd, fill = prior_reward)) +
  facet_wrap2(~ categorical_age, strip=strip) +
  geom_bar(stat = "identity", position = "dodge", colour="black") +
  geom_errorbar(aes(ymin = action1TowardsPrevEnd - se, ymax = action1TowardsPrevEnd + se), 
                position = position_dodge(width = 0.9), width = 0.25) +
  labs(title = "Empirical data",
       x = "Non-traversal reward",
       y = "Prob. of choosing associated island") +
  scale_fill_manual(name = "Non-traversal reward", values=c("#7ECAC9","#325379")) +
  scale_x_discrete(guide = guide_axis(n.dodge = 2)) +
  ylim(0,.7) +
  theme_classic(base_size = 12) + 
  theme(axis.text.x=element_text(size=11))

summary_data <- data.frame(
  action1TowardsPrevEnd = c(0.5, 0.5),
  prior_reward = c("No Reward", "Reward")
)
p2 <- ggplot(summary_data, aes(x = prior_reward, y = action1TowardsPrevEnd, fill = prior_reward)) +
  geom_bar(stat = "identity", position = "dodge", colour="black") +
  labs(title = "Model-free prediction",
    x = "Non-traversal reward",
       y = "Prob. of choosing associated island") +
  ylim(0,.7) +
  scale_fill_manual(name = "Non-traversal reward", values=c("#7ECAC9","#325379")) +
  theme_classic(base_size = 12) + 
  theme(axis.text.x=element_text(size=11)) + 
  theme(legend.position = "none")

combined_plot <- p2 + plot_spacer() + p1 + plot_layout(ncol = 3, width=c(1,.2,3))
ggsave("figures/combined_plot_mf.png", combined_plot, width = 11, height = 11/3)

summary_data <- summarySE(data, measurevar="action1TowardsPrevEnd",
                        groupvars=c("prior_shop_choice","prior_reward","sub","categorical_age"))
summary_data <- summarySE(data, measurevar="action1TowardsPrevEnd",
                        groupvars=c("prior_shop_choice","prior_reward","categorical_age"))

p1 <- ggplot(summary_data, aes(x = prior_shop_choice, y = action1TowardsPrevEnd, fill = prior_reward)) +
  geom_bar(stat = "identity", position = "dodge", colour="black") +
  geom_errorbar(aes(ymin = action1TowardsPrevEnd - se, ymax = action1TowardsPrevEnd + se), 
                position = position_dodge(width = 0.9), width = 0.25) +
  labs(title = "Empirical data",
       x = "Previous shop choice on island",
       y = "Prob. of choosing associated island") +
  scale_fill_manual(name = "Non-traversal reward", values=c("#7ECAC9","#325379")) +
  scale_x_discrete(guide = guide_axis(n.dodge = 2), limits=rev) +
  facet_wrap2(~ categorical_age, strip=strip) +
  theme_classic(base_size = 12) + 
  theme(axis.text.x=element_text(size=11)) 

summary_data <- data.frame(
  prior_shop_choice = c("Observed", "Observed", "Neighboring", "Neighboring"),
  action1TowardsPrevEnd = c(0.25, 0.666, .5, .5),
  prior_reward = c("No Reward", "Reward", "No Reward", "Reward")
)
summary_data$prior_shop_choice <- factor(summary_data$prior_shop_choice,
                                         levels = c("Observed", "Neighboring"))

p2 <- ggplot(summary_data, aes(x = prior_shop_choice, y = action1TowardsPrevEnd, fill = prior_reward)) +
  geom_bar(stat = "identity", position = "dodge", colour="black") +
  labs(title = "SR-based prediction",
    x = "Previous shop choice on island",
       y = "Prob. of choosing associated island") +
  ylim(0,.7) +
  scale_fill_manual(name = "Non-traversal reward", values=c("#7ECAC9","#325379")) +
  theme_classic(base_size = 12) + 
  theme(axis.text.x=element_text(size=11)) + 
  theme(legend.position = "none")

combined_plot <- p2 + plot_spacer() + p1 + plot_layout(ncol = 3, width=c(1,.2,3))
ggsave("figures/combined_plot_sr.png", combined_plot, width = 11, height = 11/3)

```

``` {r plot sr by age}
summary_data <- summarySE(data, measurevar="action1TowardsPrevEnd",
                        groupvars=c("prior_shop_choice","prior_reward","sub","categorical_age"))
summary_data <- summarySE(data, measurevar="action1TowardsPrevEnd",
                        groupvars=c("prior_shop_choice","prior_reward","categorical_age"))

ggplot(summary_data, aes(x = prior_shop_choice, y = action1TowardsPrevEnd, fill = prior_reward)) +
  geom_bar(stat = "identity", position = "dodge", colour="black") +
  geom_errorbar(aes(ymin = action1TowardsPrevEnd - se, ymax = action1TowardsPrevEnd + se), 
                position = position_dodge(width = 0.9), width = 0.25) +
  labs(title = "Successor Representation effect by age group",
       x = "Previous shop choice on island",
       y = "Prob. of choosing associated island") +
  scale_fill_manual(name = "Non-traversal reward", values=c("#7ECAC9","#325379")) +
  scale_x_discrete(guide = guide_axis(n.dodge = 2), limits=rev) +
  facet_wrap2(~ categorical_age, strip=strip) +
  theme_classic(base_size = 12) + 
  theme(axis.text.x=element_text(size=11), aspect.ratio=9/5) 
ggsave("figures/experiment3_sr.png")
```
