---
title: "Experiment 1"
output: html_document
date: "2024-07-23"
---

## Participants across age leverage structured knowledge to update their behavior

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
library(readr)
library(ggplot2)
library(afex)
library(Rmisc)
library(dplyr)
library(sjPlot)
library(tidyr)
library(ggh4x)
library(pROC)
```

### Loading Data
All continuous variables are z-scored
```{r read data, echo=T, results='hide', warning=FALSE}
scale_this <- function(x){
  (x - mean(x, na.rm=TRUE)) / sd(x, na.rm=TRUE)
}

learning_data <- read_csv("../data/experiment1_rest/preprocessed/learning_data.csv",show_col_types = FALSE)
learning_data$age_z <- scale_this(learning_data$age)
learning_data$trial_z <- scale_this(learning_data$trial_num)
learning_data$block_order <- scale_this(learning_data$block_order)

relearning_data <- read_csv("../data/experiment1_rest/preprocessed/relearning_data.csv",show_col_types = FALSE)
relearning_data$age_z <- scale_this(relearning_data$age)
relearning_data$trial_z <- scale_this(relearning_data$trial)
relearning_data$block_order <- scale_this(relearning_data$block_order)

data <- read_csv("../data/experiment1_rest/preprocessed/data.csv",show_col_types = FALSE)
data$age_z <- scale_this(data$age)
data$block_order <- scale_this(data$block_order)
data$second_stage_test_z <- scale_this(data$second_stage_test)

data_filtered_secondstage_accuracy <- data %>%
  group_by(subject_id) %>%
  filter(all(second_stage_test >= 0.75)) %>%
  ungroup()

first_stage_test_data <- read_csv("../data/experiment1_rest/preprocessed/first_stage_test.csv",show_col_types = FALSE)
first_stage_test_data$age_z <- scale_this(first_stage_test_data$age)
first_stage_test_data$trial_z <- scale_this(first_stage_test_data$trial_num)
first_stage_test_data_trial0 <- first_stage_test_data %>%
  filter(trial_num == 0)

exclusions_data <- read_csv("../data/experiment1_rest/preprocessed/excluded.csv", show_col_types = FALSE)
```

```{r define model functions, include=FALSE}
my_tab_model <- function(model){
  tab_model(model, 
          df.method = "wald",
          show.stat=TRUE,
          CSS = list(css.tdata = 'padding:0.1cm;padding-left:0.2cm;padding-right:0.2cm;'),
          transform = NULL,
          string.stat = "z",
          string.p = "p > |z|",
          show.re.var = FALSE,
          show.icc = FALSE,
          show.r2 = FALSE,
          wrap.labels = 100,
          show.ngroups = FALSE,
          show.obs = FALSE)
}

my_plot_model <- function(model){
  plot_model(model, 
          show.intercept = T,
          value.offset = 0.35,
          transform = NULL,
          show.values = T,) + 
  geom_hline(yintercept = 0, linetype = "dashed", alpha = .5)
}
```

### Exclusion Criteria
Participants were excluded for >20 browser interactions, >15% timeouts on learning or memory trials, >4 comprehension errors, >5 catch trials missed, or failing to learn to criterion in the learning phase of the task
```{r exclusions, warning=FALSE}
criteria <- list(
  broswer_interact = function(df) df$interactions > 20,
  timeouts = function(df) (df$task_timeouts > 0.15 | df$memory_timeouts > 0.15),
  comprehension_error = function(df) df$comprehension_retries > 4,
  catch_missed = function(df) df$catch_correct < 11,
  learning_accuracy = function(df) (df$block1_second_half_accuracy < 0.75 | df$block2_second_half_accuracy < 0.75),
  dots_missed = function(df) df$dots_missed > 4,
  parental_interference = function(df) df$parental_interference
)

# Initialize results data frame
criteria_counts <- data.frame(
  Criteria = character(),
  Total = integer(),
  Children = integer(),
  Adolescents = integer(),
  Adults = integer(),
  stringsAsFactors = FALSE
)

# Make a copy of the data
df <- exclusions_data

# Loop through criteria
for (i in seq_along(criteria)) {
  criterion_name <- names(criteria)[i]
  criterion <- criteria[[i]]
  
  # Apply criterion
  flagged <- criterion(df)
  
  # Total count
  total_count <- sum(flagged)
  
  # Count per age group
  age_counts <- table(df$categorical_age[flagged])
  
  # Get counts or set to 0 if missing
  children_count <- ifelse("Children" %in% names(age_counts), age_counts["Children"], 0)
  adolescents_count <- ifelse("Adolescents" %in% names(age_counts), age_counts["Adolescents"], 0)
  adults_count <- ifelse("Adults" %in% names(age_counts), age_counts["Adults"], 0)
  
  # Append to result
  criteria_counts <- rbind(
    criteria_counts,
    data.frame(
      Criteria = criterion_name,
      Total = total_count,
      Children = children_count,
      Adolescents = adolescents_count,
      Adults = adults_count,
      stringsAsFactors = FALSE
    )
  )
  
  # Exclude flagged participants
  df <- df[!flagged, ]
}

# View result
criteria_counts
```

### Learning: Did participants learn to make optimal choices over the course of the learning blocks?
Looking at optimal two-stage choices across the learning phase and average accuracy in the last 10 learning trials.
```{r learning model 1}
learning_mod <- glmer(correct ~ age_z * trial_z * block_condition + (trial_z * block_condition | subject_id), 
                  data = learning_data,
                  family = binomial,
                  contrasts = list(block_condition="contr.sum"))

my_tab_model(learning_mod)

# Look at average accuracy in last 10 trials of learning accross age
summarySE(data, measurevar="last_10_accuracy", groupvars=c("categorical_age"))
```

```{r learning plot}
ld <- summarySE(learning_data, measurevar="correct", groupvars=c("categorical_age","trial_bin"))
ld$categorical_age <- factor(ld$categorical_age, levels = c('Children', 'Adolescents', 'Adults'))

ggplot(ld, aes(x=trial_bin, y=correct, color=categorical_age)) + 
    geom_point() +
    geom_line(linewidth=0.8) +
    geom_errorbar(aes(ymin=correct-se, ymax=correct+se), width=.7) +
    xlab("Trial Number") +
    ylab("Proportion of Optimal Choices") +
    scale_color_manual(name="Age Group", values=c("#B27EE0","#7c44ad", "#501b80")) +
    theme_classic(base_size = 18) +
    theme(aspect.ratio=1, legend.position = c(0.7, 0.3), legend.box.background=element_rect(),legend.box.margin=margin(5,5,5,5)) 
ggsave("figures/experiment1_learning.png", height=4)
```

### Re-learning: Did participants learn to make optimal choices during relearning?
```{r relearning mod}
relearning_mod <- glmer(correct ~ age_z * trial_z * block_condition + (trial_z * block_condition | subject_id),
                  data= relearning_data,
                  family = binomial,
                  contrasts = list(block_condition="contr.sum"))
my_tab_model(relearning_mod)

relearning_mod_block_order <- glmer(correct ~ age_z * trial_z * block_condition * block_order + (trial_z * block_condition | subject_id),
                  data= relearning_data,
                  family = binomial,
                  contrasts = list(block_condition="contr.sum"))
my_tab_model(relearning_mod_block_order)
```

```{r relearning plot}
relearning_data$trial_bin <- relearning_data$trial %/% 2 * 2
sub_means <- summarySE(relearning_data, measurevar="correct", groupvars=c("block_condition","trial_bin","subject_id","categorical_age"))
means <- summarySE(sub_means, measurevar="correct",
                        groupvars=c("block_condition","trial_bin","categorical_age"))
means$categorical_age <- factor(means$categorical_age, levels = c('Children', 'Adolescents', 'Adults'))
means <- means %>% mutate(block_condition=recode(block_condition,
`original`="Control",
`revaluation`="Revaluation"))

strip <- strip_themed(background_x = elem_list_rect(fill = c("#F8746B", "#10BEC3")))
ggplot(means, aes(y=correct,x=trial_bin, color=categorical_age)) +
  geom_point() +
  geom_line(linewidth=0.8) +
  scale_color_manual(name="categorical_age", values=c("#B27EE0","#7c44ad", "#501b80")) +
  geom_errorbar(aes(ymin=correct-se, ymax=correct+se), width=.7)+
  facet_wrap2(~ block_condition, scales = "free", strip=strip) +
  ylab("Proportion of Optimal Choices") + 
  scale_y_continuous(limits = c(0.5, 1)) +
  theme_classic(base_size = 18) +
  theme(aspect.ratio=2,legend.position="none") + 
  xlab("Trial Number") 
ggsave("figures/experiment1_relearning.png",height = 4)
```

### Second-stage test: did participants learn from direct experience in re-learning?
```{r second stage test model}
summarySE(data, measurevar="second_stage_test", groupvars=c("block_condition"))

summarySE(data, measurevar="second_stage_test", groupvars=c("categorical_age"))

second_stage_test_mod <- lmer(second_stage_test ~ age_z * block_condition + (1 | subject_id),
                   data=data,
                   contrasts = list(block_condition="contr.sum"))
my_tab_model(second_stage_test_mod)
```

```{r 2nd stage test plot}
data$categorical_age <- factor(data$categorical_age, levels = c('Children', 'Adolescents', 'Adults'))

# Calculate means and SEs
means <- summarySE(data, measurevar="second_stage_test", groupvars=c("categorical_age", "block_condition"))

# Themed strip for facet
strip <- strip_themed(
  background_x = elem_list_rect(fill = c("#B27EE0","#7c44ad", "#501b80")),
  text_x = list(element_text(colour = "white")),
  background_y = elem_list_rect(fill = c("#C1ED5C"))
)

ggplot(data, aes(x=second_stage_test, y=block_condition, fill=block_condition)) +
  geom_bar(stat="summary", fun="mean", position=position_dodge(width=0.7), width=1, color="black") +
  geom_point(position=position_jitter(width=0, height=0.15), alpha=0.5, shape=21, color="black") +
  geom_errorbar(
    data = means,
    aes(
      x = second_stage_test,
      y = block_condition,
      xmin = second_stage_test - se,
      xmax = second_stage_test + se
    ),
    width=0.5,
    position = position_dodge(width = 0.7),
    inherit.aes = FALSE
  ) +
  facet_wrap2('categorical_age', strip=strip, ncol=1) +
  scale_x_continuous(limits = c(0, 1.0), breaks = c(0, 0.25, 0.5, 0.75, 1)) +
  xlab("Second Stage Test Accuracy") +
  ylab("Block Condition") +
  theme_classic(base_size = 16) + 
  scale_fill_manual(values=c("#F8746B", "#10BEC3")) +
  theme(axis.text.y=element_text(size=10), legend.position="none", aspect.ratio=1/3)

ggsave("figures/experiment1_secondstagetest.png", height = 4)
```

### First-stage test: did participants update their first-stage choices?
```{r revaluation 1}
reval_mod <- lmer(reval_score ~  age_z * block_condition + (1 | subject_id),
                  data= data,
                  contrasts = list(block_condition="contr.sum"))
my_tab_model(reval_mod)

reval_mod_by_trial <- glmer(one_trial_revaluation ~  age_z * block_condition * trial_z + (trial_z | subject_id),
                  data= first_stage_test_data,
                  family = binomial,
                  contrasts = list(block_condition="contr.sum"))
my_tab_model(reval_mod_by_trial)
```

```{r first stage plot}
data$categorical_age <- factor(data$categorical_age, levels = c('Children', 'Adolescents', 'Adults'))

# Calculate means and SEs
means <- summarySE(data, measurevar="accuracy", groupvars=c("categorical_age", "block_condition"))

# Themed strip for facet
strip <- strip_themed(
  background_x = elem_list_rect(fill = c("#B27EE0","#7c44ad", "#501b80")),
  text_x = list(element_text(colour = "white")),
  background_y = elem_list_rect(fill = c("#C1ED5C"))
)

ggplot(data, aes(x=accuracy, y=block_condition, fill=block_condition)) +
  geom_bar(stat="summary", fun="mean", position=position_dodge(width=0.7), width=1, color="black") +
  geom_point(position=position_jitter(width=0, height=0.15), alpha=0.5, shape=21, color="black") +
  geom_errorbar(
    data = means,
    aes(
      x = accuracy,
      y = block_condition,
      xmin = accuracy - se,
      xmax = accuracy + se
    ),
    width=0.5,
    position = position_dodge(width = 0.7),
    inherit.aes = FALSE
  ) +
  facet_wrap2('categorical_age', strip=strip, ncol=1) +
  scale_x_continuous(limits = c(0, 1.0), breaks = c(0, 0.25, 0.5, 0.75, 1)) +
  xlab("First Stage Test Accuracy") +
  ylab("Block Condition") +
  theme_classic(base_size = 16) + 
  scale_fill_manual(values=c("#F8746B", "#10BEC3")) +
  theme(axis.text.y=element_text(size=10), legend.position="none", aspect.ratio=1/3)

ggsave("figures/experiment1_firststagetest.png", height = 4)

sub_means <- summarySE(first_stage_test_data, measurevar="one_trial_revaluation", groupvars=c("block_condition","trial_num","subject_id","categorical_age"))
means <- summarySE(sub_means, measurevar="one_trial_revaluation",
                        groupvars=c("block_condition","trial_num","categorical_age"))
means$categorical_age <- factor(means$categorical_age, levels = c('Children', 'Adolescents', 'Adults'))
means <- means %>% mutate(block_condition=recode(block_condition,
`original`="Control",
`revaluation`="Revaluation"))

strip <- strip_themed(background_x = elem_list_rect(fill = c("#F8746B", "#10BEC3")))
ggplot(means, aes(y=one_trial_revaluation,x=trial_num, color=categorical_age)) +
  geom_point() +
  geom_line(linewidth=0.8) +
  scale_color_manual(name="categorical_age", values=c("#B27EE0","#7c44ad", "#501b80")) +
  geom_errorbar(aes(ymin=one_trial_revaluation-se, ymax=one_trial_revaluation+se), width=.7)+
  facet_wrap2(~ block_condition, scales = "free", strip=strip) +
  ylab("First-stage choice updating") + 
  scale_y_continuous(limits = c(0, .6)) +
  scale_x_continuous(labels = function(x) x + 1) +
  theme_classic(base_size = 18) +
  theme(aspect.ratio=2,legend.position="none") + 
  xlab("Trial Number") 
ggsave("figures/experiment1_firststagetest_trial.png", height = 4)
```

``` {r reval plot}
data$categorical_age <- factor(data$categorical_age, levels = c('Children', 'Adolescents', 'Adults'))
means <- summarySE(data, measurevar="reval_score", groupvars=c("categorical_age","block_condition"))
strip <- strip_themed(background_x = elem_list_rect(fill = c("#B27EE0","#7c44ad", "#501b80")),text_x=list(element_text(colour = "white")), background_y = elem_list_rect(fill = c("#C1ED5C")))
ggplot(data, aes(x=block_condition, y=reval_score, fill =block_condition, color =block_condition)) +
    geom_hline(yintercept=0) +
    facet_wrap2('categorical_age', strip=strip) +
    geom_line(aes(group = subject_id), color = 'black', linewidth = .1, alpha = .5) +
    geom_point(alpha = .5, position = position_jitter(width = 0.12)) +
    geom_line(aes(x = block_condition, y = reval_score, group = categorical_age), color = "black", size = 1, data = means) +
    geom_point(aes(x = block_condition, y = reval_score), size = 5, alpha = 1, data=means) +
    scale_y_continuous(limits = c(-0.25, 1.01)) +
    xlab("Block Condition") +
    ylab("Replanning Score") +
    theme_classic(base_size = 16) + 
    scale_color_manual(values=c("#F8746B", "#10BEC3")) +
    theme(axis.text.x=element_text(size=12),legend.position="none", aspect.ratio=9/6)
ggsave("figures/experiment1_revaluation.png")
```


## Supplemental Models

### Looking at relation between second stage test accuracy and replanning 
``` {r excluding by second stage test}
reval_mod_secondstage_accuracy <- lmer(reval_score ~  age_z * block_condition * second_stage_test_z + (1 | subject_id),
                  data= data,
                  contrasts = list(block_condition="contr.sum"))
my_tab_model(reval_mod_secondstage_accuracy)

reval_mod_filtered_secondstage_accuracy <- lmer(reval_score ~  age_z * block_condition + (1 | subject_id),
                   data=data_filtered_secondstage_accuracy,
                   contrasts = list(block_condition="contr.sum"))
my_tab_model(reval_mod_filtered_secondstage_accuracy)

excluded_subjects <- data %>%
  group_by(subject_id) %>%
  filter(any(second_stage_test < 0.75)) %>%
  distinct(subject_id)

data <- data %>%
  mutate(filtered_for_second_stage = subject_id %in% excluded_subjects$subject_id)

data$filtered_for_second_stage <- factor(
    data$filtered_for_second_stage, 
    levels = c(TRUE, FALSE) # TRUE comes first
)

age_demographics_excluded <- data %>%
  filter(filtered_for_second_stage == TRUE) %>%
  group_by(categorical_age) %>%
  summarise(n = n()/2)
age_demographics_excluded

labs <- c("TRUE" = "Second Stage Accuracy <75%", 
          "FALSE" = "Second Stage Accuracy >=75%")

means <- summarySE(data, measurevar="reval_score", groupvars=c("filtered_for_second_stage","block_condition"))
strip <- strip_themed(background_x = elem_list_rect(fill = c("#B27EE0","#7c44ad", "#501b80")),text_x=list(element_text(colour = "white")), background_y = elem_list_rect(fill = c("#C1ED5C")))
ggplot(data, aes(x=block_condition, y=reval_score, fill =block_condition, color =block_condition)) +
    geom_hline(yintercept=0) +
    facet_wrap2('filtered_for_second_stage', strip=strip, labeller = labeller(filtered_for_second_stage = labs)) +
    geom_line(aes(group = subject_id), color = 'black', linewidth = .1, alpha = .5) +
    geom_point(alpha = .5, position = position_jitter(width = 0.12)) +
    geom_line(aes(x = block_condition, y = reval_score, group = filtered_for_second_stage), color = "black", size = 1, data = means) +
    geom_point(aes(x = block_condition, y = reval_score), size = 5, alpha = 1, data=means) +
    scale_y_continuous(limits = c(-0.25, 1.01)) +
    xlab("Block Condition") +
    ylab("Replanning Score") +
    theme_classic(base_size = 16) + 
    scale_color_manual(values=c("#F8746B", "#10BEC3")) +
    theme(axis.text.x=element_text(size=14),legend.position="none", aspect.ratio=1.2) 

ggsave("figures/experiment1_filtered_secondstagetest.png")
```


### Memory: does memory for first-stage stimuli relate to replanning behavior
``` {r memory models}
memory_data <- read_csv("../data/experiment1_rest/preprocessed/memory_data.csv",show_col_types = FALSE)
confidence.map <- c("definitely_new"=1, "maybe_new"=2, "maybe_old"=3, "definitely_old"=4)   
memory_data$confidence_rating <- confidence.map[as.character(memory_data$response0)]
memory_AUCs <- memory_data %>%
select(subject_id, environment, ground_truth, confidence_rating) %>%
    group_by(subject_id, environment) %>%
    mutate(memory_auc = as.numeric(roc(ground_truth ~ confidence_rating)$auc)) %>%
    select(subject_id, environment, memory_auc) %>%
    unique()
memory_data <- merge(data, memory_AUCs, by=c("subject_id","environment")) 

memory_data$reval_score <- scale_this(memory_data$reval_score)

memory_mod <- lmer(memory_auc ~  age_z * block_order * block_condition * reval_score + environment + (1|subject_id),
                   data=memory_data, 
                   contrasts = list(block_condition="contr.sum",environment="contr.sum"))
my_tab_model(memory_mod)
```